---
title: "Data Products - Project Pitch"
author: "BB"
date: "08/11/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r libraries_data, warning=FALSE,message=FALSE,echo=FALSE}

library(readxl)
library(dplyr)
library(openxlsx)
library(ggplot2)
library(broom)
library(plotly)


raw <- read_excel("UKCovidData.xlsx", sheet = 5)
pop <- read_excel("UKPopData.xlsx", sheet = 19)

# Population Data Processing

popMale <- as.numeric(pop[8,4:94])
popFemale <- as.numeric(pop[8,96:186])
popTotal <- popMale + popFemale
popProp <- popTotal / 66796807
popPropGrouped <- c(
  sum(popProp[1]),
  sum(popProp[2:5]),
  sum(popProp[6:10]),
  sum(popProp[11:15]),
  sum(popProp[16:20]),
  sum(popProp[21:25]),
  sum(popProp[26:30]),
  sum(popProp[31:35]),
  sum(popProp[36:40]),
  sum(popProp[41:45]),
  sum(popProp[46:50]),
  sum(popProp[51:55]),
  sum(popProp[56:60]),
  sum(popProp[61:65]),
  sum(popProp[66:70]),
  sum(popProp[71:75]),
  sum(popProp[76:80]),
  sum(popProp[81:85]),
  sum(popProp[86:90]),
  sum(popProp[91]))

# Death Data Processing

deathsDF <- data.frame(
  total=t(raw[8,3:44]),
  fiveYearAverage=t(raw[10,3:44]),
  covidExcess=t(raw[8,3:44])-t(raw[10,3:44]),
  covidOfficial=t(raw[18,3:44])
)

dFrame=list()
sumDeaths=list()
dFrame[[1]] <- data.frame(raw[43:62,3:44]) #Male
dFrame[[2]] <- data.frame(raw[65:84,3:44]) #Female
dFrame[[3]] <- data.frame(raw[21:40,3:44]) #Both

for(i in 1:3) {
  rownames(dFrame[[i]])<-pull(raw[21:40,2]) #Naming age groups
  sumDeaths[[i]] <- rowSums(dFrame[[i]]) #sumDeaths[[1(Male),2(Female),3(Both)]]
}

dates<-as.Date(as.numeric(raw[5,3:44]),origin = "1899-12-30")
dF<-dFrame[[1]]

ages_all<-c(0:90)
AgesMed <- c(0.5,3,7.5,12.5,17.5,22.5,27.5,32.5,37.5,42.5,47.5,52.5,57.5,62.5,67.5,72.5,77.5,82.5,87.5,90)
dF$AgesMed <- AgesMed
sD <- data.frame(SumDeaths = sumDeaths[[1]], AgesMed = AgesMed)

names(dF)[1:42] <- paste("Week", as.character(c(1:42)),sep='')

dF <- dF[c(-1,-20),]
sD <- sD[c(-1,-20),]

ageModel<-lm(SumDeaths~poly(AgesMed,3),data = sD)

aA <- data.frame(AgesAll=ages_all)

agePredictions<-predict(ageModel,data.frame(AgesMed=ages_all))

## Death Distribution Pre-Covid19

preCovidDeaths <- data.frame(raw[21:40,3:12]) #Data covers 03/01/2020 to 06/03/2020

preCovidMeans <- rowMeans(preCovidDeaths)

preCovidDist <- preCovidMeans/sum(preCovidMeans)

## Death Distribution During Covid19

midCovidDeaths <- data.frame(raw[21:40,17:20])

midCovidMeans <- rowMeans(midCovidDeaths)

midCovidDist <- midCovidMeans/sum(midCovidMeans)

## Quick Comparative Plot

qComp <- data.frame(x=AgesMed,preC=preCovidDist,midC=midCovidDist)

cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
## Excess Covid Deaths (per Week)

# absExcessDeaths <- midCovidDeaths - preCovidDeaths # - Not sure if needed anymore

excessDeathsMeans <- midCovidMeans - preCovidMeans

excessDeathsPercentage <- (midCovidMeans/preCovidMeans)*100-100 #Percentage Difference in death rate per age group

excessDeathsProp <- excessDeathsMeans / sum(excessDeathsMeans) #Sum of excess death means is greater than average covid confirmed deaths

# Modeling the Death Distribution

covidDeathAgeData <- data.frame(x=AgesMed,excDP=excessDeathsProp,preCD=preCovidDist)

covidDeathAgeModel <- lm(excDP[-c(1:8)]~poly(x[-c(1:8)],5),data=covidDeathAgeData)

sSplineObj<-smooth.spline(covidDeathAgeData$x,covidDeathAgeData$excDP,df=6)


# Model Plot - Altered model (for p<1/10000, p=0)

modelPDF <- predict(sSplineObj,data.frame(x=ages_all))

modelPDF <- data.frame(x=modelPDF[[1]]$x,y=modelPDF[[2]]$x)

modelPDF_original <- modelPDF

for(i in 1:dim(modelPDF)[1]) if(modelPDF$y[i]<0) modelPDF$y[i]=0

modelPDF$y<-modelPDF$y/sum(modelPDF$y)

```

## Introduction

An exact record of how many people per age group died from Covid-19 in England and Wales over the past year has not been made publicly available. 

What has been made available are two sets of information:

- Total amount of official Covid deaths per week.
- Total deaths in the country per week broken down by age group. 

The following model predicts the amount of people who died of Covid-19 per age group.

## Introduction

By comparing how the proportions of deaths changed per age group before and during Covid-19. Performing regression analysis on the resulting proportion provides a model of how many people died for a specific age.

The difference between the excess deaths and official deaths illustrates a rough sense of how many died from Covid-19 unofficialy, i.e. People that were never tested for the virus.

## Proportions Pre and Post Covid-19

```{r cars, echo = FALSE}

ggplot(qComp) + geom_line(aes(x=x,y=preC,color="Pre Covid-19")) + geom_line(aes(x=x,y=midC,color="Mid Covid-19")) + scale_fill_manual(values=cbPalette) + ylab("Death Probability") + xlab("Age")

```

## Spline Regression Driving Model

```{r pressure, warning=FALSE,message=FALSE,echo=FALSE}

ggplot(modelPDF_original) + geom_line(aes(x=x,y=y,color="Mid Covid-19 Model")) + scale_fill_manual(values=cbPalette) + ylab("Death Probability") + xlab("Age")

```